\documentclass[11pt]{article}

\input{sections/preamble.tex}
\begin{document}

\maketitle

\begin{abstract}
  \input{sections/0_abstract.tex}
\end{abstract}

\input{sections/1_intro.tex}
\input{sections/2_derivation.tex}
\input{sections/3_backward.tex}
\input{sections/4_linear_tent_profile.tex}
\input{sections/5_micro.tex}
\input{sections/6_numerics.tex}
\input{sections/7_results.tex}
\input{sections/8_conclusions.tex}
\input{sections/9_code.tex}
% ------------------------------------------------------------
\begin{thebibliography}{99}
% ------------------------------------------------------------

\bibitem{Elf2005}
J.~Elf, G.-W.~Li, and X.~S.~Xie,
``Probing transcription factor dynamics at the single-molecule level in a
living cell,''
\textit{Science} \textbf{316} (2007), 1191--1194.

\bibitem{Metzler2014}
R.~Metzler, J.-H.~Jeon, A.~G.~Cherstvy, and E.~Barkai,
``Anomalous diffusion models and their properties: non-stationarity,
non-ergodicity, and ageing at the centenary of single particle tracking,''
\textit{Phys.\ Chem.\ Chem.\ Phys.} \textbf{16} (2014), 24128--24164.

\bibitem{Peleg2018}
O.~Peleg, J.~Peters, M.~Salcedo, and L.~Mahadevan,
``Collective mechanical adaptation of honeybee swarms,''
\textit{Nature Physics} \textbf{14} (2018), 1193--1198.

\bibitem{Ocko2014}
S.~A.~Ocko and L.~Mahadevan,
``Collective thermoregulation in bee clusters,''
\textit{J.~R.~Soc.\ Interface} \textbf{11} (2014), 20131033.

\bibitem{Schuss2015}
Z.~Schuss,
\textit{Brownian Dynamics at Boundaries and Interfaces: In Physics, Chemistry,
and Biology},
Applied Mathematical Sciences 186, Springer, 2015.

\bibitem{Benichou2010}
O.~B\'enichou, C.~Chevalier, J.~Klafter, B.~Meyer, and R.~Voituriez,
``Geometry-controlled kinetics,''
\textit{Nature Chemistry} \textbf{2} (2010), 472--477.

\bibitem{Redner2001}
S.~Redner,
\textit{A Guide to First-Passage Processes},
Cambridge University Press, 2001.

\bibitem{Gardiner2009}
C.~W. Gardiner,
\textit{Stochastic Methods: A Handbook for the Natural and Social Sciences},
4th ed., Springer, 2009.

\bibitem{Pavliotis2014}
G.~A. Pavliotis,
\textit{Stochastic Processes and Applications: Diffusion Processes, the
Fokker--Planck and Langevin Equations},
Texts in Applied Mathematics 60, Springer, 2014.

\bibitem{Kloeden1992}
P.~E. Kloeden and E.~Platen,
\textit{Numerical Solution of Stochastic Differential Equations},
Stochastic Modelling and Applied Probability 23, Springer, 1992.

\bibitem{Risken1996}
H.~Risken,
\textit{The Fokker--Planck Equation: Methods of Solution and Applications},
2nd ed., Springer Series in Synergetics 18, Springer, 1996.

\bibitem{Benichou2011}
O.~B\'enichou, C.~Loverdo, M.~Moreau, and R.~Voituriez,
``Intermittent search strategies,''
\textit{Rev.\ Mod.\ Phys.} \textbf{83} (2011), 81--129.

\bibitem{Metzler2000}
R.~Metzler and J.~Klafter,
``The random walk's guide to anomalous diffusion: a fractional dynamics
approach,''
\textit{Physics Reports} \textbf{339} (2000), 1--77.

\bibitem{Bressloff2013}
P.~C.~Bressloff and J.~M.~Newby,
``Stochastic models of intracellular transport,''
\textit{Rev.\ Mod.\ Phys.} \textbf{85} (2013), 135--196.

\bibitem{LeVeque2007}
R.~J.~LeVeque,
\textit{Finite Difference Methods for Ordinary and Partial Differential
Equations: Steady-State and Time-Dependent Problems},
SIAM, 2007.

\bibitem{Iserles2009}
A.~Iserles,
\textit{A First Course in the Numerical Analysis of Differential Equations},
2nd ed., Cambridge University Press, 2009.

\bibitem{Evans2010}
L.~C.~Evans,
\textit{Partial Differential Equations},
2nd ed., Graduate Studies in Mathematics 19, AMS, 2010.

\end{thebibliography}

\end{document}
