
% ------------------------------------------------------------
\section{Backward equation, flux-jump condition, and MFPT}
\label{sec:backward}
% ------------------------------------------------------------

Let $\tau=\inf\{t\ge 0:\text{absorbed}\}$ be the (random) first passage time and
let $T(x)=\mathbb{E}[\tau\mid x_0=x]$ be the MFPT.

\subsection{Derivation from the survival probability}

Define the survival probability $S(x,t)=\Pr(\tau>t\mid x_0=x)$.  A standard
calculation in stochastic processes~\cite{Gardiner2009,Pavliotis2014} shows that
$S$ satisfies the backward equation
\begin{equation}
\partial_t S \;=\; \mathcal{L}^\dagger S, \qquad S(x,0)=1,\quad S(x,\infty)=0,
\label{eq:survival-eq}
\end{equation}
where $\mathcal{L}^\dagger$ is the adjoint of the forward generator.
From the integral representation $T(x)=\int_0^\infty S(x,t)\,dt$, integrating
\eqref{eq:survival-eq} in $t$ from $0$ to $\infty$,
\[
\mathcal{L}^\dagger T
=\mathcal{L}^\dagger\!\int_0^\infty\!S\,dt
=\int_0^\infty\!\partial_t S\,dt
=S(x,\infty)-S(x,0)
=-1.
\]
For Fick's form the operator $\mathcal{L}$ is self-adjoint
(Section~\ref{sec:derivation}), so $\mathcal{L}^\dagger=\mathcal{L}$ and the
MFPT satisfies
\begin{equation}
\boxed{\;
\frac{d}{dx}\!\left[D(x)\,\frac{dT}{dx}\right] \;=\; -1,
\qquad x\in(-\pi,\pi)\setminus\{0\},
\;}
\label{eq:backwardODE}
\end{equation}
subject to periodic boundary conditions $T(-\pi)=T(\pi)$,
$T'(-\pi)=T'(\pi)$ and the trap condition derived below.

\subsection{Flux-jump condition at the trap}

Integrating the forward equation \eqref{eq:forward} on $[-\varepsilon,\varepsilon]$ and
letting $\varepsilon\to 0^+$,
\[
\frac{d}{dt}\!\int_{-\varepsilon}^{\varepsilon}\!p\,dx
= \Bigl[D(x)\partial_x p\Bigr]_{-\varepsilon}^{\varepsilon}
-\kappa\int_{-\varepsilon}^{\varepsilon}\!\delta(x)p\,dx,
\]
the left-hand side vanishes as $\varepsilon\to 0$ (by boundedness of $p$) and the
$\delta$-integral picks up $p(0,t)$, yielding the \emph{flux-jump condition} for
the forward problem:
\begin{equation}
D(0^+)\,\partial_x p\big|_{0^+}
-D(0^-)\,\partial_x p\big|_{0^-}
=\kappa\,p(0,t).
\label{eq:jump-forward}
\end{equation}
Taking the adjoint of this condition yields the corresponding condition for the
MFPT.  Because $D$ and $T$ are continuous at $x=0$ and, by the symmetry of
$D(x)$, $T$ is an even function of $x$ so that $T'(0^-)=-T'(0^+)$, the adjoint
flux-jump condition simplifies to the \emph{Robin-like} condition
\begin{equation}
\boxed{\;
2\,D(0)\,T'(0^+) \;=\; \kappa\,T(0).
\;}\label{eq:robin}
\end{equation}

\subsection{Closed-form solution of the backward equation}

Equation~\eqref{eq:backwardODE} can be integrated directly.  On $(0,\pi)$,
\[
(D(x)T'(x))' \;=\; -1
\quad\Longrightarrow\quad
D(x)T'(x) \;=\; -x + C
\quad\Longrightarrow\quad
T'(x)=\frac{C-x}{D(x)},
\]
for an integration constant $C$.  By the symmetry $T(-x)=T(x)$ the same $C$
appears on $(-\pi,0)$.  The constant $C$ is fixed by the periodic boundary
condition $T'(\pi)=T'(-\pi)$.  Since by symmetry $T'(-\pi)=-T'(\pi)$, we need
$T'(\pi)=0$, hence $D(\pi)\cdot 0 = C-\pi$, giving
\begin{equation}
C \;=\; \pi.
\label{eq:C-pi}
\end{equation}
Integrating $T'(x)=(\pi-x)/D(x)$ from $0$ to $x>0$,
\begin{equation}
\boxed{\;
T(x) \;=\; T(0) \;+\; \int_0^{|x|}\!\frac{\pi-y}{D(y)}\,dy.
\;}\label{eq:T-of-x}
\end{equation}

It remains to determine $T(0)$.  From \eqref{eq:robin}, $T(0)=2D(0)T'(0^+)/\kappa
=2\pi/\kappa$ using $D(0)T'(0^+)=C=\pi$.  Alternatively, integrating
\eqref{eq:backwardODE} over the whole ring produces the same result as follows.
The periodic contribution of $(DT')'$ vanishes, and the only surviving
contribution comes from the flux jump at $x=0$, which by \eqref{eq:jump-forward}
extracts $-\kappa T(0)$.  Hence
\[
-\kappa\,T(0) \;=\; \int_{-\pi}^{\pi}\!(-1)\,dx = -2\pi
\quad\Longrightarrow\quad
T(0) \;=\; \frac{2\pi}{\kappa}.
\]

\begin{remark}
The two identities $C=\pi$ and $T(0)=2\pi/\kappa$ are consistent: they are
respectively a local statement (flux conservation at the outermost boundary of
the ring) and a global statement (integrated source matches absorbed flux).
Both are independent of the profile $D(x)$.
\end{remark}

Putting it together,
\begin{equation}
\boxed{\;
T(x) \;=\; \frac{2\pi}{\kappa} \;+\; \int_0^{|x|}\!\frac{\pi-y}{D(y)}\,dy.
\;}\label{eq:T-final}
\end{equation}
The spatial average over $[-\pi,\pi]$, using Fubini to exchange the order of
integration in the region $\{0\le y\le x\le\pi\}$, is
\begin{align}
\langle T\rangle
&=\frac{1}{2\pi}\int_{-\pi}^\pi T(x)\,dx
=\frac{2\pi}{\kappa}
+\frac{1}{\pi}\int_0^\pi\!\int_0^x\!\frac{\pi-y}{D(y)}\,dy\,dx \notag\\
&=\frac{2\pi}{\kappa}
+\frac{1}{\pi}\int_0^\pi (\pi-y)\cdot\frac{\pi-y}{D(y)}\,dy
=\frac{2\pi}{\kappa}
+\frac{1}{\pi}\int_0^\pi\!\frac{(\pi-y)^2}{D(y)}\,dy.
\label{eq:T-avg-general}
\end{align}
Equation~\eqref{eq:T-avg-general} is the central analytical result of this
report.  It exhibits an additive decomposition
\[
\langle T\rangle
=\underbrace{\frac{2\pi}{\kappa}}_{\text{dwell time at the trap}}
+\underbrace{\frac{1}{\pi}\int_0^\pi\!\frac{(\pi-y)^2}{D(y)}\,dy}_{\text{travel time on the ring}}.
\]
With $\kappa$ constant the two contributions are independent; the spatial
profile $D(x)$ affects only the travel term.  We return to the implications of
this decomposition in Section~\ref{sec:kappa-stories}.
